---
title: "Bike Benefits"
author: "David F. Severski"
date: "July 9, 2016"
output: 
  html_document: 
    code_folding: hide
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background and Motivation

While in the factory store of our local bean-to-bar chocolatier, Theo's 
Chocolate, the spouse pointed out a sign that bicyclists could purchase a sticker 
for their helmet which would give them 50% off one of Theo's standard bars 
each time they came in. I've long wanted a discount program for local folks 
coming in so I jumped on this chance. Looking at the sticker, it turned out not 
to be something specific to Theo's but a larger program run by [Bicycle 
Benefits](http://www.bicyclebenefits.org/). What a terrific idea!

An initial glance at the hosted map showed a large number of participating 
business  nearby as well as a long list of cities where member business could be 
found. The map is a standard Leaflet module with the data available via a 
simple GET request or two. This stirred my curiosity to explore the data.

# Data Acquisiton and Prepping

A standard set of libraries here. I'm trying to use the `purrr` library more 
to build familarity with its powerful fuctional idioms.

```{r load_libraries, message=FALSE, warning=FALSE}
library(jsonlite)
library(purrr)
library(httr)
library(dplyr)
library(DT)
library(ggplot2)
library(scales)
```

Fetch and clean the data.

```{r acquisitons_inc}
# fetch all cities
cities_url <- 'http://www.bicyclebenefits.org/search/cities'
#cities <- fromJSON(content(GET(cities_url), as = "text"), flatten = TRUE)
#save(cities, file = "data/bikebenefits_cities.Rdata")
load("data/bikebenefits_Cities.Rdata")

# fetch the details for a given city
get_biz <- function(city_id, base_url = 'http://www.bicyclebenefits.org/search/members?city_id=')  {
  search_url <- paste0(base_url, city_id)
  dat <- fromJSON(content(GET(search_url), as = "text"), flatten = TRUE)
  dat$members$city_id <- city_id
  dat$city$city_id <- city_id
  return(dat)
}

#dat <- map(cities$id, get_biz)
#save(dat, file = "data/bikebenefits.Rdata")
load("data/bikebenefits.Rdata")

# coerce member business into a dataframe and peel off categories into 
# a more normalized table
members <- map(dat, "members") %>% compact %>% bind_rows() %>% 
  filter(!is.na(id))
categories <- members %>% distinct(category.id, category.name, category.logo)
categories[categories$category.logo == "","category.logo"] <- NA
members <- members %>% select(-c(category.name, category.logo))

# why do so many places publish the email addresses of their participants?!
members <- select(members, -email)

# coerce the states into a dataframe
states <- map(dat, "city")
states <- do.call(rbind, states) %>% as_data_frame
states$state_name <- map_chr(states$state, "name")
states <- select(states, -state) %>% simplify_all %>% bind_rows
```

# Exploration

Three quick glances of the data at the level of geographic distribution, 
member business types, and business specifically located in Seattle.

## Geography
```{r exploration}
members %>% count(city_id, sort = TRUE) %>% 
  left_join(cities, by = c("city_id" = "id")) %>% 
  select(city_name = name, state.name, n) %>% 
  datatable(rownames = NA)
```

Lots of activity in Charlotte. How does this look at a 
state level?

```{r}
members %>% count(city_id, sort = TRUE) %>% 
  left_join(cities, by = c("city_id" = "id")) %>% 
  select(city_name = name, state.name, n) %>% 
  group_by(state.name) %>% tally(n, sort = TRUE) %>% 
  datatable(rownames = NA)
```

Wow. Wisconsin has a lot going on, while most of Washington's activity 
is limited to Seattle.

## Business Type

Produce a table of the relative distribution of member categories per city.

```{r}
type_breakdown <- members %>% group_by(city_id) %>% 
  mutate(city_count = n()) %>% ungroup %>% 
  group_by(city_id, category.id) %>% 
  summarize(total_per_city = n(),
            cat_pct = n() / max(city_count)) %>% ungroup %>% 
  arrange(city_id, desc(cat_pct)) %>% 
  left_join(cities, by = c("city_id" = "id")) %>% 
  left_join(categories, by = c("category.id" = "category.id"))
type_breakdown %>% 
  select(name, category.name, total_per_city, cat_pct) %>% 
  datatable(rownames = FALSE) %>% formatPercentage('cat_pct')
```

## Seattle Uber Alles

Finally, a quick view of the Seattle specific data.

```{r}
members %>% filter(city_id == '3') %>% 
  left_join(categories, by = c("category.id" = "category.id")) %>% 
  select(-c(starts_with("shipping"), extra_stickers, id, category.id, city_id, category.logo, status)) %>% 
  datatable(rownames = FALSE)
```

Show the relative distributions of the types of member business for Seattle 
and Portland. This may be the only time I've ever used sportball colors in a 
plot. I feel a little dirty.

```{r seattle_v_portland}
facet_labels <- c(
  '3' = "Seattle, WA",
  '23' = "Portland, OR")
type_breakdown %>% filter(city_id == '3' | city_id == '23') %>% 
  mutate(city_id = as.factor(city_id)) %>% 
  left_join(categories, by = c("category.id" = "category.id", "category.name" = "category.name")) %>% 
  ggplot(., aes(x = category.name, y = cat_pct)) -> gg
gg <- gg + geom_bar(stat = "identity", aes(fill = city_id))
gg <- gg + scale_fill_manual(values = c("#002244", "#EAE727"), guide = FALSE)
gg <- gg + labs(title = "Distribution of Bikeshare Participants by Category",
                subtitle = "Seattle vs. Portland",
                caption = "Source bikeshare.org\nRetrieved July 9, 2016",
                x = "Business Type",
                y = element_blank())
gg <- gg + scale_y_continuous(labels = percent)
gg <- gg + theme_minimal()
gg <- gg + coord_flip()
gg <- gg + facet_grid(. ~ city_id, labeller = as_labeller(facet_labels))
gg
```
